---
title: "Data Tidying"
author: "Emily"
date: "5/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
```

```{r}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_df)
```

Data are in wide format. All and notes columns could be problematic. 

```{r}
unique(catch_df$notesRegCode)
```

This information should really be in metadata; won't be used for analysis. 

```{r}
catch_df %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>% 
  head()
```

This isn't saved to a new object yet. Just checking to make sure that this is what we want. 

```{r}
catch_df <- catch_df %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  head(catch_df)
```

Selecting columns by name is more powerful; accounts for changes to DF that may change row/column numbers. 

```{r}
catch_df <- catch_df %>% 
  gather(key = Species, value = Catch, -Region, -Year)

head(catch_df)
```

```{r}
catch_df %>% 
  spread(key=Species, value=Catch) 
```

Goes back to a wide frame. 

```{r}
catch_df <- catch_df %>% 
  rename(catch_thousands = Catch)

head(catch_df)
```

```{r}
catch_df %>% 
  mutate(catch = catch_thousands*1000)
```

Error because catch_thousands is a character vector. 

```{r}
test <- as.integer(catch_df$catch_thousands)
which(is.na(test==TRUE))
```

```{r}
catch_df[401, ]
```

```{r}
View(catch_df)
```

```{r}
catch_df <- catch_df %>% 
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
  mutate(catch_thousands = as.integer(catch_thousands)) %>% 
  mutate(catch = catch_thousands*1000)

```

```{r}
head(catch_df, 30)
```

#Analysis

```{r}
catch_df %>% 
  group_by(Species) %>% 
  summarize(mean_catch = mean(catch))
```

```{r}
catch_df %>% 
  group_by(Region) %>%
  summarize(n = n())
```

```{r}
catch_df %>% 
  filter(Region == "SSE") %>% 
  head()
```

```{r}
catch_df %>% 
  group_by(Region) %>% 
  summarise(mean_catch = mean(catch, na.rm=TRUE)) %>% 
  arrange(desc(mean_catch))
```

